Measuring Perceptions and Preferences for Meritocracy

### 2021-10-07 paper-blinded.Rmd
version final sometida a SJR, es paper-blinded, y la información de autores se traspasa a title-page

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization. Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141). Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

@@

@@

@@

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy? The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny. Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery. It consists of a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts. Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ or what it _should_ be?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs cover almost anything related to subjective factors. To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while actually referring to perceptions (i.e. what is). This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

@@

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic. Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.

@@

Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly (i.e. no error or residual variance) by the selected indicators, going as far as to propose that "... In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654]. Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, somewhat surprisingly they finally chose to build a sum index despite proving a multidimensional latent structure.

## Conceptual proposal for studying meritocracy {-#instrumentprop}

@@

```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
 knitr::include_graphics("input/images/Fig1.generalf.png")
```

@@

There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme. However, it is important to mention that the items were not originally designed as a specific measure of meritocracy. The variables are presented below organized in perceptions and preferences:

@@

First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable. Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first-order latent variables parental attributes, racial and gender background, and networks.

@@

Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models. The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)). Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification. Attending to these constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp). In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
 knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
```

Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions. Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum. Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$). From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct. All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.

@@

The items of the meritocracy scale were identical to the original. The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig:appmod)).

@@

Table \@ref(tab:fcloads-inv) shows the results of the CFA estimation of the meritocracy scale in this new sample. As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.

@@

Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation. When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal. Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012]. Still, the comparability of latent means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$. Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007]. Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.

@@

The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale open several avenues for future research. For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred. Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies. The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions. Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research. Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.

@@

# (APPENDIX) Appendix {-}

@@

## Appendix study 1

@@

## Appendix study 3

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and operationalization of meritocracy beliefs. Based on this critical analysis, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel scale designed with this purpose (N=2,141). Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

@@

<!---
\pagebreak

@@

--->

@@

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy? The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny. Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above. Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs use to cover almost anything related to subjective factors. To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while actually referring to perceptions (i.e. what is). This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

@@

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic. Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.

@@

Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly (i.e. no error or residual variance) by the selected indicators, going as far as to propose that "... In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654]. Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.

## A conceptual proposal for studying meritocracy {-#instrumentprop}

@@

```{r merit-model, echo=FALSE, fig.cap = "Conceptual Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
 knitr::include_graphics("input/images/Fig1.generalf.png")
```

@@

There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):

@@

Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models. The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)). Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.

@@

Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp). In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
 knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
```

Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions. Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum. Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$). From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct. All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing and testing new scale for measuring meritocracy.

@@

The items of the meritocracy scale were identical to the original. The items were administrated in a randomized order (which corresponds to the application modality of Group 3 as showed in Figure \@ref(fig:appmod)).

@@

Table \@ref(tab:fcloads-inv) in the Appendix shows the results of the CFA estimation of the meritocracy scale in this new sample. As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.

@@

Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation. When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal. Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012]. Still, the comparability of latent means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$. Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007]. Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.

@@

The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research. For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred. Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies. The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions. Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research. Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.

# Transparency statement

This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in an public repository in Github. This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.

@@

# Appendix {-}

@@

**Study 1**

@@

**Study 3**

### 2021-10-07 paper.Rmd
version final sometida a SJR, es paper-blinded, y la información de autores se traspasa a title-page

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization. Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141). Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization. Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141). Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---
